---
title: "Module 3 - Visualization and EDA- Part 3"
author: "Matthew Lawlor"
date: "10/12/2020"
output: github_document
---

```{r setup, include=FALSE}
library (tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
  
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

##Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

weather_df
```

## `group_by`

```{r}
weather_df %>% 
  group_by(name, month)
  ## ungroup(month) removes group structure
```

## counting things

count month observations

```{r}
weather_df %>%
  group_by(name, month) %>% 
  summarize(n_obs = n())
```

we can use `count`

```{r}
weather_df %>% 
  count(name, month, name = "n_obs")
```

**Never** use base R's `table`

```{r, eval = FALSE}
weather_df %>% 
  pull(month) %>% 
  table()
```

other helpful counters

```{r}
weather_df %>%
  group_by(name, month) %>% 
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

## A digression on 2x2 tables

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% ##filter != things not equal to Waikiki
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE      ~ ""
    )
  ) %>% 
  group_by (name, cold) %>% 
  summarize(count = n())
  ## could turn this into 2x2 table using pivot_wider
```

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% ##filter != things not equal to Waikiki
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE      ~ ""
    )
  ) %>% 
  janitor::tabyl(name, cold)  ##note this is a dataframe, as opposed to base R "table" function
```

## General summaries

You can do lots of summaries.

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  )
```

This is a dataframe! So, we can make a plot.

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  ) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()

```

Suppose you want to summarize many columns.

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean)
  )
```

Reminder: sometimes your results are easier to read in another format

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)
```

## `group_by` and `mutate`

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```

What about window functions

ranking...
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  mutate(
    temp_rank = min_rank(tmax)) %>% ##can also use desc() to reverse order of min_rank
  filter(temp_rank == 1)
```

lag: moves variables down a row...

```{r}
weather_df %>% 
  group_by(name) %>%
  mutate(
    ##lag_temp = lag(tmax),
    temp_change = tmax - lag(tmax)
    ) %>% 
  summarize(
    temp_change_max = max(temp_change, na.rm = TRUE),
    temp_change_sd = sd(temp_change, na.rm = TRUE)
)
```


## Quick note

summarize only gets you so far. Single number output

