---
title: "Module 3 - Visualization and EDA"
author: Matthew Lawlor
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

##Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatterplots!!

Create my first scatterplot ever

```{r}
ggplot(weather_df, aes(x= tmin, y=tmax)) + 
  geom_point()
```

New approach, same plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save and edit plot objet

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x= tmin, y = tmax))

weather_plot + geom_point()
```

## Advanced scatterplot

Start with the same one and make it fancy

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

What about the "aes" placement?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth()
```

Let's facet some things

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .2) + ##alpha refers to density of dots
  geom_smooth(se = FALSE, size = 2) +
  facet_grid(. ~ name) ##facet_grid(row ~ column),"."  means don't create
```

Let's combine some elements and try a new plot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se=FALSE) +
  facet_grid(. ~ name)
```

## Some small notes

How many geoms have to exist?
What we are saying is you can have whatever geoms you want

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se=FALSE)
```  

Here we have got the smooth geometries without the point geomet

You can use a neat geom!

```{r}

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d()
```

##Univariate plots

Histograms are really great

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Can we add color...
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge") + ##dodge important if not creating different facets
  facet_grid(. ~ name) ## kinda makes things less clear here
```

Let's try a new geometry

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .4, adjust = .5) #alpha for color density, adjust for smoothing
```

What about boxplots

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

Other univariate plot types

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = .5) + 
  ##geom_violin=boxplot with some aspect of density plot
  stat_summary(fun=median)
```

Ridge plots -- the most popular plot of 2017

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges() ##pulls apart multiple variables in a density plot
```

